---
title: "Visualization Part 2"
output: html_notebook
---

## A working example

The previous notebook was a simple introduction to the syntax for constructing a `ggplot` object.  This notebook will introduce a more meaningful example using landcover data from the NSF National Ecological Observation Network (NEON) project.  For more information you can look at the [NEON website](http://www.neonscience.org)

Load the tidyverse library and ggrepel

```{r libraries}
library(tidyverse)
library(ggrepel)
```


Dr. Montana gave us a sample vegetation structure data from the San Juaquin Experimental Range in California.

n = 362 records

### Load the data

```{r load-structure-data}
sjer_veg <- read_csv("data/D17_2013_SJER_vegStr.csv")
# View(sjer_veg)
sjer_veg
```



Instead of summarizing by plot (as in the example), I summarized by taxonid. 

```{r maxcan_summ}
sjer_veg %>% 
  group_by(taxonid) %>% 
  summarise(m = mean(maxcanopydiam)) 
 
```



Created a bar chart based on the taxonid grouping, and summarizing the mean for the percent living canopy within each of the 11 taxonid groups

```{r summary-viz}
sjer_veg %>% 
 
  group_by(taxonid) %>% 
  summarise(l = mean(livingcanopy)) %>% 
  ggplot(aes(x = taxonid, y = l)) +
    geom_col() + # a column plot for each entry
    ggtitle("SJER Distribution of Mean Living Canopy by Taxon Id") +
    xlab("Taxon ID") + 
    ylab("Living Canopy (%)") +
    theme_minimal() + #change the theme
    theme(plot.margin=unit(c(.5,1,.5,1), "cm"), #adjust the margin (top, right, bottom, left)
          axis.text.x = element_text(angle = -60, hjust = -0.1)) # rotate the axis labels

```


 So now I created a new data frame taxgroups, in which I summarize the mean values for three variables, as grouped by taxonid


=
```{r group_by_livcan}
taxgroups <- sjer_veg %>% 
  select(taxonid, livingcanopy, maxcanopydiam, stemheight) %>% 
  group_by(taxonid) %>%
  summarise(livingcanopy = mean(livingcanopy), maxcanopydiam = mean(maxcanopydiam), stemheight = mean(stemheight))
 
taxgroups

```

Great!  Now I want to create an indicator of how plentiful each tree type is within each taxonid in this dataset (n = 362)  

the frequncy ranged from 1 to 123

```{r count_taxonNum}

taxonum <- sjer_veg %>% 
  select(taxonid) %>% 
  count(taxonid)

taxonum

```


so now I wanted to join the two dataframes (taxgroups and taxonum) I created by taxonid into a new dataframe called allgroups



```{r joinum}

allgroups <- full_join(taxgroups, taxonum, by = "taxonid")

allgroups

```

Plotting allgroups, we have stemheight and maxcanopy on the x,y axis, with the living canopy a color scale, and the frequency of taxoid occurances used as a scale for the size of the group.

Found a good example to label the points using a package called ggrepel, that worked nicely.  Thankyou to "Tung" on stackoverflow. 
Retrieved on April 4, 2018, from https://stackoverflow.com/questions/15624656/label-points-in-gem-point 


```{r plot_allgroups}

allgroups %>% 
  ggplot(aes(x = stemheight, 
             y = maxcanopydiam, 
             label = taxonid, size = n)) + 
    geom_point(aes(color = livingcanopy)) +
    scale_color_gradient(low = "red", high = "green") +
  geom_label_repel(,
                   box.padding = 0.35, 
                   point.padding = 0.5,
                   segment.color = 'grey50', 
                   size = 2.5) +
                   theme_classic(base_size = 12) +
   ggtitle("Distribution of 
           Mean Stem Height and Max Canopy Diameter by Taxon Id, 
              with Mean Living Canopy Percentage (color scale)
              and Taxon Id occurance frequency (point size)") +
    xlab("Stem Height (m)") + 
    ylab("Max Canopy Diameter (m)") 

```




